---
title: "DATA_CHECK_HW1"
author: "Vishal"
date: "2025-01-27"
output: pdf_document
---

# DSC 480: Initial Tasks# 1. Computational Setup- Set up IDEs, GitHub repos, forked, and cloned the repository to the local machine.

## 2. Summary Statistics and Exploratory Analysis, calculate summary statistics for the 'events' and 'report' tables. 

this is going to help us understand the distribution of data

-   we gotta import the data from 'read_clean.R'

```{r}
events <- read_csv("events_toy.csv")
report <- read_csv("report_toy.csv")
```

-   now we gotta compute the summary statistics for both of the datasets

    ```{r}
    head(summary(events))
    head(summary(report))
    ```

    -   cleaning the data( took this from clean data)

```{r}
events_full <- events %>% # Converting Timestamps to Date Format
  mutate_at(vars(contains("timestamp")), ~as.POSIXct(.x, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate_at(vars(contains("timestamp")), ~as_date(.x)) %>% # Renaming Columns
  rename_with(~str_replace(., "timestamp", "date"), contains("timestamp")) %>%
  mutate(across(event_type, ~ str_replace(.x, " ", "_"))) |> # Cleaning event_type Strings
  mutate(across(event_type, ~as.factor(.x))) #Converting event_type to Factor


report_full <- report %>%
  mutate_at(vars(contains("timestamp")), ~as.POSIXct(.x, format = "%Y-%m-%d %H:%M:%S")) %>% # Converting Timestamps to Date Format
  mutate_at(vars(contains("timestamp")), ~as_date(.x)) %>% #Renaming Columns
  rename_with(~str_replace(., "timestamp", "date"), contains("timestamp")) %>%
    mutate(across(event_type, ~ str_replace(.x, " ", "_"))) |> #Cleaning event_type Strings
  mutate(across(event_type, ~as.factor(.x))) #Converting event_type to Factor

sum(is.na(events_full$event_start_date)) # check if any rows have NA values due to invalid timestamp formats
```

print out the summary of the cleaned data

```{r}
summary(events_full)
```

now for the other file

```{r}
summary(report_full)
```

```{r}
summary(events_full$event_start_date) 
summary(report_full$event_start_date)
```

some visualizations

```{r}
ggplot(events_full, aes(x = event_type)) +        # Sets up the ggplot with events_full dataset and maps  the event_type to the x axis
  geom_bar(fill = "lightblue") +                 # Creates a bar plot with bars filled in light blue
  theme_minimal() +                              # Applies a minimal theme
  labs(                                          # Add labels for the title, x-axis and y-axis
    title = "Count of Events by Type",
    x = "Event Type",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels by 45° for better readability
```

```{r}
ggplot(events_full, aes(x = coverage_duration_days)) +  # Maps coverage_duration_days to the x-axis
  geom_histogram(                                       # Adds a histogram 
    binwidth = 50,                                      # Sets the bin width to group the data into intervals of 50 days
    fill = "orange",                                # Fills the bars with  orange color
    color = "black"                                     # Outlines  the bars in black
  ) +
  theme_minimal() +                                     # Apply aminimal theme
  labs(                                                 # Add labels to the title, x-axis and y-axis
    title = "Distribution of Coverage Duration Days",
    x = "Coverage Duration (Days)",
    y = "Count"
  )
```

```{r}
ggplot(events_full, aes(x = coverage_duration_days, y = event_duration_days)) +  # Maps coverage_duration_days to xaxis and event_duration_days to yaxis
  geom_point(alpha = 0.4, fill = "red", color = "lightgreen") +  # Add points with some transparency (alpha),(this helps with overlapping so you can see better and all the other points), and  we set color to green
  theme_minimal() +                              # Applies a minimal theme
  labs(                                          # Add labels for the title, x-axis, and y-axis
    title = "Coverage Duration vs. Event Duration",
    x = "Coverage Duration (Days)",
    y = "Event Duration (Days)"
  )
```

```{r}
# Filter for Ibis and Pre-Ibis events
ibis_events <- events_full %>% filter(ibis_observation == 1)  # Events during Ibis
pre_ibis_events <- events_full %>% filter(non_ibis_observation == 1)  # Events before Ibis

head(ibis_events)
head(pre_ibis_events)

pre_ibis_admissions_count <- events_full %>%
  filter(event_start_date <= ibis_coverage_start_date) %>%  # Filter for events before Ibis coverage started
  filter(event_start_date >= coverage_start_date) %>%       # Ensure events are after the general coverage start
  select(pid, event_type, event_start_date) %>%             # Select relevant columns
  filter(event_type == "inpatient") %>%                     # Filter for inpatient events
  group_by(pid) %>%                                         # Group by patient ID
  summarize(admit_count = n()) %>%                          # Count the number of admissions per patient
  right_join(patients, by = "pid") %>%                      # Include all patients, even those with zero admissions
  mutate(admit_count = ifelse(is.na(admit_count), 0, admit_count))  # Replace NA counts with 0

head(pre_ibis_admissions_count)

```

```{r}

## 3. Check Erik's Values
### Task
Compare the calculated summary statistics and exploratory results to the Google Doc values to ensure accuracy. 

Relevant comparisons:
- Total admissions on Ibis vs. Pre-Ibis.
- Coverage days on Ibis vs. Pre-Ibis.
- Recreate the admissions count bar plots.

---

## 4. Descriptive Statistics for Combined Groups
### Goals
Perform descriptive statistics for the `Unicare-study` group using Ibis and compare it to the combined dataset when Ibis is not used (MIIA).

### Key Metrics
1. Total days under observation.
2. Total inpatient days.
3. Proportion of patients with at least one inpatient admission.
4. Counts of inpatient admissions.

---

## 5. Data Cleaning Function
### Goal
Create a function that reads and cleans data, using the logic in `read_clean.R`.

read_clean_data <- function(file) {
  # Add code here based on `read_clean.R`
}
```
