---
title: "DATA_CHECK_HW1"
author: "Vishal"
date: "2025-01-27"
output: pdf_document
---

```{r}
# Load libraries
library(tidyverse)  
```

# DSC 480: Initial Tasks# 1. Computational Setup- Set up IDEs, GitHub repos, forked, and cloned the repository to the local machine.

## 2. Summary Statistics and Exploratory Analysis, calculate summary statistics for the 'events' and 'report' tables. 

this is going to help us understand the distribution of data

-   we gotta import the data from 'read_clean.R'

```{r}
setwd("/Users/vishmurda/DSC480_GitHub_Repository")
getwd

```

```{r}
events <- read_csv("/Users/vishmurda/DSC480_GitHub_Repository/events_toy.csv")
report <- read_csv("/Users/vishmurda/DSC480_GitHub_Repository/report_toy.csv")
```

-   now we gotta compute the summary statistics for both of the datasets

    ```{r}
    head(summary(events))
    head(summary(report))
    ```

    -   cleaning the data( took this from clean data)

```{r}
events_full <- events %>% # Converting Timestamps to Date Format
  mutate_at(vars(contains("timestamp")), ~as.POSIXct(.x, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate_at(vars(contains("timestamp")), ~as_date(.x)) %>% # Renaming Columns
  rename_with(~str_replace(., "timestamp", "date"), contains("timestamp")) %>%
  mutate(across(event_type, ~ str_replace(.x, " ", "_"))) |> # Cleaning event_type Strings
  mutate(across(event_type, ~as.factor(.x))) #Converting event_type to Factor


report_full <- report %>%
  mutate_at(vars(contains("timestamp")), ~as.POSIXct(.x, format = "%Y-%m-%d %H:%M:%S")) %>% # Converting Timestamps to Date Format
  mutate_at(vars(contains("timestamp")), ~as_date(.x)) %>% #Renaming Columns
  rename_with(~str_replace(., "timestamp", "date"), contains("timestamp")) %>%
    mutate(across(event_type, ~ str_replace(.x, " ", "_"))) |> #Cleaning event_type Strings
  mutate(across(event_type, ~as.factor(.x))) #Converting event_type to Factor

sum(is.na(events_full$event_start_date)) # check if any rows have NA values due to invalid timestamp formats
```

print out the summary of the cleaned data

```{r}
summary(events_full)
```

now for the other file

```{r}
summary(report_full)
```

```{r}
summary(events_full$event_start_date) 
summary(report_full$event_start_date)
```

some visualizations

```{r}
ggplot(events_full, aes(x = event_type)) +        # Sets up the ggplot with events_full dataset and maps  the event_type to the x axis
  geom_bar(fill = "lightblue") +                 # Creates a bar plot with bars filled in light blue
  theme_minimal() +                              # Applies a minimal theme
  labs(                                          # Add labels for the title, x-axis and y-axis
    title = "Count of Events by Type",
    x = "Event Type",
    y = "Count"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))  # Rotate x-axis labels by 45° for better readability
```

```{r}
ggplot(events_full, aes(x = coverage_duration_days)) +  # Maps coverage_duration_days to the x-axis
  geom_histogram(                                       # Adds a histogram 
    binwidth = 50,                                      # Sets the bin width to group the data into intervals of 50 days
    fill = "orange",                                # Fills the bars with  orange color
    color = "black"                                     # Outlines  the bars in black
  ) +
  theme_minimal() +                                     # Apply aminimal theme
  labs(                                                 # Add labels to the title, x-axis and y-axis
    title = "Distribution of Coverage Duration Days",
    x = "Coverage Duration (Days)",
    y = "Count"
  )
```

```{r}
ggplot(events_full, aes(x = coverage_duration_days, y = event_duration_days)) +  # Maps coverage_duration_days to xaxis and event_duration_days to yaxis
  geom_point(alpha = 0.4, fill = "red", color = "lightgreen") +  # Add points with some transparency (alpha),(this helps with overlapping so you can see better and all the other points), and  we set color to green
  theme_minimal() +                              # Applies a minimal theme
  labs(                                          # Add labels for the title, x-axis, and y-axis
    title = "Coverage Duration vs. Event Duration",
    x = "Coverage Duration (Days)",
    y = "Event Duration (Days)"
  )
```

```{r}
# Filter for Ibis and Pre-Ibis events
ibis_events <- events_full %>% filter(ibis_observation == 1)  # Events during Ibis
pre_ibis_events <- events_full %>% filter(non_ibis_observation == 1)  # Events before Ibis

head(ibis_events)
head(pre_ibis_events)

pre_ibis_admissions_count <- events_full %>%
  filter(event_start_date <= ibis_coverage_start_date) %>%  # Filter for events before Ibis coverage started
  filter(event_start_date >= coverage_start_date) %>%       # Ensure events are after the general coverage start
  select(pid, event_type, event_start_date) %>%             # Select relevant columns
  filter(event_type == "inpatient") %>%                     # Filter for inpatient events
  group_by(pid) %>%                                         # Group by patient ID
  summarize(admit_count = n()) %>%                          # Count the number of admissions per patient
  right_join(patients, by = "pid") %>%                      # Include all patients, even those with zero admissions
  mutate(admit_count = ifelse(is.na(admit_count), 0, admit_count))  # Replace NA counts with 0

pre_ibis_admissions_count

post_ibis_admissions_count <- events_full %>%
  filter(event_start_date > ibis_coverage_start_date) %>%   # Filter for events after Ibis coverage started
  filter(event_start_date <= ibis_coverage_end_date) %>%    # Ensure events are within the Ibis coverage period
  select(pid, event_type, event_start_date) %>%             # Select relevant columns
  filter(event_type == "inpatient") %>%                     # Filter for inpatient events
  group_by(pid) %>%                                         # Group by patient ID
  summarize(admit_count = n()) %>%                          # Count the number of admissions per patient
  right_join(patients, by = "pid") %>%                      # Include all patients, even those with zero admissions
  mutate(admit_count = ifelse(is.na(admit_count), 0, admit_count))  # Replace NA counts with 0

post_ibis_admissions_count

str(events_full)
summary(events_full$event_start_date)
summary(events_full$ibis_coverage_start_date)
summary(events_full$ibis_coverage_end_date)


#pre-ibis admissions for unicare group
pre_ibis_admissions_count %>% filter(org_name == "Unicare - Study") %>% 
    summarize(unicare_pre_admits  = sum(admit_count))

#total pre-Ibis admissions for all groups
pre_ibis_admissions_count %>%
  summarize(total_pre_admits_all = sum(admit_count, na.rm = TRUE))


post_ibis_admissions_count %>% filter(org_name == "Unicare - Study") %>%
    summarize(unicare_post_admits = sum(admit_count))

post_ibis_admissions_count %>%
  summarize(total_post_admits_all = sum(admit_count, na.rm = TRUE))



# Combines the results into a single tibble
admissions_summary <- tibble(
  Group = c("Unicare - Study", "All Groups"),
  Pre_Ibis_Admissions = c(
    pre_ibis_admissions_count %>% filter(org_name == "Unicare - Study") %>% summarize(sum(admit_count)) %>% pull(),
    pre_ibis_admissions_count %>% summarize(sum(admit_count, na.rm = TRUE)) %>% pull()
  ),
  Post_Ibis_Admissions = c(
    post_ibis_admissions_count %>% filter(org_name == "Unicare - Study") %>% summarize(sum(admit_count)) %>% pull(),
    post_ibis_admissions_count %>% summarize(sum(admit_count, na.rm = TRUE)) %>% pull()
  )
)

print(admissions_summary)


```

```{r}
# Reshape data for plotting
admissions_summary_long <- admissions_summary %>%  # Start with the admissions_summary table
  pivot_longer(                                   # Use pivot_longer to reshape the data from wide to long format
    cols = c(Pre_Ibis_Admissions, Post_Ibis_Admissions), # Specify the columns to "gather" (pre- and post-Ibis admissions)
    names_to = "Period",                         # Create a new column named "Period" to store the column names
    values_to = "Admissions"                     # Create a new column named "Admissions" to store the corresponding values
  )

# Plot
ggplot(admissions_summary_long, aes(x = Group, y = Admissions, fill = Period)) +  # Create a ggplot object
  geom_bar(stat = "identity", position = "dodge") +  # Add a bar plot with identity stat (use raw values) and "dodge" to create side-by-side bars
  theme_minimal() +                                  # Apply a minimal theme for a clean look
  labs(                                              # Add labels to the plot
    title = "Pre- and Post-Ibis Admissions",         # Title of the plot
    x = "Group",                                     # Label for the x-axis
    y = "Admissions Count",                          # Label for the y-axis
    fill = "Period"                                  # Legend title for the fill (colors) representing "Period"
  )
```

```{r}
admissions_summary <- admissions_summary %>%
  mutate(
    Percent_Change = ((Post_Ibis_Admissions - Pre_Ibis_Admissions) / Pre_Ibis_Admissions) * 100
  )

print(admissions_summary)
```

now we got to do the same thing but for coverage on preibis vs post ibis

```{r}
####### 


# Pre- and Post-Ibis Coverage Days for Unicare - Study
report_full %>% 
  filter(ibis_coverage_start_date >= coverage_start_date) %>%  # Filter rows where Ibis coverage started after or on the general coverage start date
  mutate(pre_ibis_covered_days_check = ibis_coverage_start_date - coverage_start_date) %>%  # Calculate the days covered before Ibis started
  mutate(post_ibis_covered_days_check = ibis_coverage_end_date - ibis_coverage_start_date) %>%  # Calculate the days covered during Ibis
  select(pid, org_name, pre_ibis_covered_days_check, post_ibis_covered_days_check) %>%  # Select relevant columns for the calculation
  filter(org_name == "Unicare - Study") %>%  # Filter data to include only the Unicare - Study group
  summarize(
    unicare_pre_total = sum(pre_ibis_covered_days_check),  # Sum up the total pre-Ibis coverage days for Unicare members
    unicare_post_total = sum(post_ibis_covered_days_check)  # Sum up the total post-Ibis coverage days for Unicare members
  )

# Calculate Pre- and Post-Ibis Coverage Days for All Members
report_full %>% 
  filter(ibis_coverage_start_date >= coverage_start_date) %>%  # Filter rows where Ibis coverage started after or on the general coverage start date
  mutate(pre_ibis_covered_days_check = ibis_coverage_start_date - coverage_start_date) %>%  # Calculate pre-Ibis coverage days
  mutate(post_ibis_covered_days_check = ibis_coverage_end_date - ibis_coverage_start_date) %>%  # Calculate post-Ibis coverage days
  select(pid, org_name, pre_ibis_covered_days_check, post_ibis_covered_days_check) %>%  # Select relevant columns
  summarize(
    all_member_pre_total = sum(pre_ibis_covered_days_check),  # Sum the total pre-Ibis coverage days for all members
    all_member_post_total = sum(post_ibis_covered_days_check)  # Sum the total post-Ibis coverage days for all members
  )

# Pre-Ibis coverage days for Unicare group
unicare_pre_coverage <- report_full %>%
  filter(org_name == "Unicare - Study") %>%  # Filter for Unicare group
  summarize(unicare_pre_days = sum(ibis_coverage_start_date - coverage_start_date, na.rm = TRUE))  # Calculate total pre-Ibis days

# Total Pre-Ibis coverage days for all groups
all_pre_coverage <- report_full %>%
  summarize(total_pre_days = sum(ibis_coverage_start_date - coverage_start_date, na.rm = TRUE))  # Calculate total pre-Ibis days for all groups

# Post-Ibis coverage days for Unicare group
unicare_post_coverage <- report_full %>%
  filter(org_name == "Unicare - Study") %>%  # Filter for Unicare group
  summarize(unicare_post_days = sum(ibis_coverage_end_date - ibis_coverage_start_date, na.rm = TRUE))  # Calculate total post-Ibis days

# Total Post-Ibis coverage days for all groups
all_post_coverage <- report_full %>%
  summarize(total_post_days = sum(ibis_coverage_end_date - ibis_coverage_start_date, na.rm = TRUE))  # Calculate total post-Ibis days for all groups

# Combine results into a single tibble
coverage_summary <- tibble(
  Group = c("Unicare - Study", "All Groups"),  # Define the groups
  Pre_Ibis_Coverage_Days = c(
    unicare_pre_coverage$unicare_pre_days,  # Pre-Ibis days for Unicare
    all_pre_coverage$total_pre_days         # Pre-Ibis days for all groups
  ),
  Post_Ibis_Coverage_Days = c(
    unicare_post_coverage$unicare_post_days,  # Post-Ibis days for Unicare
    all_post_coverage$total_post_days         # Post-Ibis days for all groups
  )
)

# Print the final summary table
print(coverage_summary)


# Admissions count bar plot
events_full %>%
  filter(event_type == "inpatient") %>%  # Filter for inpatient events
  mutate(period = ifelse(event_start_date <= ibis_coverage_start_date, "Pre-Ibis", "Post-Ibis")) %>%  # Define period
  group_by(org_name, period) %>%  # Group by organization and period
  summarize(admissions_count = n(), .groups = "drop") %>%  # Count admissions
  ggplot(aes(x = org_name, y = admissions_count, fill = period)) +
  geom_bar(stat = "identity", position = "dodge") +  # Create side-by-side bars
  theme_minimal() +  # Apply a minimal theme
  labs(
    title = "Admissions Count by Period",
    x = "Organization",
    y = "Admissions Count",
    fill = "Period"
      
  ) + theme(axis.text.x = element_text(angle = 45, hjust = 1))


```

```{r}


## 4. Descriptive Statistics for Combined Groups
### Goals
#Perform descriptive statistics for the `Unicare-study` group using Ibis and compare it to the combined dataset when Ibis is not used (MIIA).

### Key Metrics
#1. Total days under observation.
#2. Total inpatient days.
#3. Proportion of patients with at least one inpatient admission.
#4. Counts of inpatient admissions.

#---

## 5. Data Cleaning Function
### Goal
#Create a function that reads and cleans data, using the logic in `read_clean.R`.

#```r
#read_clean_data <- function(file) {
  # Add code here based on `read_clean.R`
#}
```
