---
title: "Tasks 2025-02-24 - Vishal"
format: pdf
---

```{r}
# Load necessary libraries
library(tidyverse)

# Read in the new events and report files
events_new <- read_csv("2025_02_18_utilization_events.csv")
report_new <- read_csv("2025_02_18_utilization_report.csv")

# Preview the data
#head(events_new)
#head(report_new)

## A bit of cleaning
events_full_new <- events_new %>%
  mutate_at(vars(contains("timestamp")), ~as.POSIXct(.x, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate_at(vars(contains("timestamp")), ~as_date(.x)) %>%
  rename_with(~str_replace(., "timestamp", "date"), contains("timestamp")) %>%
  mutate(across(event_type, ~ str_replace(.x, " ", "_"))) |>
  mutate(across(event_type, ~as.factor(.x)))


report_full_new <- report_new %>%
  mutate_at(vars(contains("timestamp")), ~as.POSIXct(.x, format = "%Y-%m-%d %H:%M:%S")) %>%
  mutate_at(vars(contains("timestamp")), ~as_date(.x)) %>%
  rename_with(~str_replace(., "timestamp", "date"), contains("timestamp")) %>%
    mutate(across(event_type, ~ str_replace(.x, " ", "_"))) |>
  mutate(across(event_type, ~as.factor(.x)))

## add comment

head(events_full_new)
head(report_full_new)
```

**Count Total Inpatient Admissions**

```{r}
# Count total inpatient admissions for Unicare-Study
pre_ibis_admissions <- events_full_new %>%
  filter(org_name == "Unicare - Study", pre_ibis_inpatient == 1, event_type == "inpatient") %>%
  summarise(total_pre_ibis_admissions = n())

during_ibis_admissions <- events_full_new %>%
  filter(org_name == "Unicare - Study", ibis_inpatient == 1, event_type == "inpatient") %>%
  summarise(total_during_ibis_admissions = n())

# Display results
pre_ibis_admissions
during_ibis_admissions
```

**Admissions Per 1000 Patient-Years**

```{r}
# Compute total patient days
patient_days <- report_full_new %>%
  summarise(
    total_days_pre_ibis = sum(covered_days_pre_ibis, na.rm = TRUE),
    total_days_on_ibis = sum(covered_days_on_ibis, na.rm = TRUE)
  )

# Convert days to years
patient_years <- patient_days %>%
  mutate(
    pre_ibis_years = total_days_pre_ibis / 365,
    on_ibis_years = total_days_on_ibis / 365
  )

# Compute admissions per 1000 patient-years
admissions_per_1000 <- tibble(
  period = c("Pre-Ibis", "During-Ibis"),
  admissions = c(pre_ibis_admissions$total_pre_ibis_admissions,
                 during_ibis_admissions$total_during_ibis_admissions),
  patient_years = c(patient_years$pre_ibis_years, patient_years$on_ibis_years)
) %>%
  mutate(admissions_per_1000_years = (admissions / patient_years) * 1000)

# Display results
admissions_per_1000
```

**Count Admissions Per Patient**

```{r}
# Count number of admissions per patient for pre-Ibis
pre_ibis_counts <- events_full_new %>%
  filter(pre_ibis_inpatient == 1, event_type == "inpatient") %>%
  group_by(pid) %>%
  summarise(admit_count = n()) %>%
  count(admit_count)

# Count number of admissions per patient for during-Ibis
during_ibis_counts <- events_full_new %>%
  filter(ibis_inpatient == 1, event_type == "inpatient") %>%
  group_by(pid) %>%
  summarise(admit_count = n()) %>%
  count(admit_count)

# Display results
pre_ibis_counts
during_ibis_counts
```

**Compute Inpatient Days Per 1000 Days**

```{r}
# Compute inpatient days per 1000 patient-days
inpatient_days_per_1000 <- tibble(
  period = c("Pre-Ibis", "During-Ibis"),
  inpatient_days = c(sum(events_full_new$pre_ibis_inpatient_days, na.rm = TRUE),
                     sum(events_full_new$ibis_inpatient_days, na.rm = TRUE)),
  total_days = c(patient_days$total_days_pre_ibis, patient_days$total_days_on_ibis)
) %>%
  mutate(inpatient_days_per_1000 = (inpatient_days / total_days) * 1000)

# Display results
inpatient_days_per_1000
```

**Statistical Tests**

```{r}
nrow(pre_ibis_counts)
nrow(during_ibis_counts)
```

```{r}
setdiff(pre_ibis_counts$pid, during_ibis_counts$pid)  # Patients missing from during_ibis_counts
setdiff(during_ibis_counts$pid, pre_ibis_counts$pid)  # Patients missing from pre_ibis_counts
```

```{r}
# Paired t-test: Compare admissions per day before and after.
t_test_result <- t.test(pre_ibis_counts$admit_count, during_ibis_counts$admit_count, paired = TRUE)
t_test_result
```

```{r}
# Wilcoxon signed-rank test: Compare admission counts.
wilcox_result <- wilcox.test(pre_ibis_counts$admit_count, during_ibis_counts$admit_count, paired = TRUE)
wilcox_result
```

```{r}
#McNemar’s Test: Check probability of admission pre vs. post.
pre_post_matrix <- matrix(c(
  sum(pre_ibis_counts$admit_count == 0 & during_ibis_counts$admit_count > 0),
  sum(pre_ibis_counts$admit_count > 0 & during_ibis_counts$admit_count == 0),
  sum(pre_ibis_counts$admit_count > 0 & during_ibis_counts$admit_count > 0),
  sum(pre_ibis_counts$admit_count == 0 & during_ibis_counts$admit_count == 0)
), nrow = 2, byrow = TRUE)

mcnemar_result <- mcnemar.test(pre_post_matrix)
mcnemar_result  
```

The McNemar test results show a statistically significant difference (p = 0.000512) in admissions before and during Ibis, meaning the intervention had a measurable impact. Since the chi-square value (12.071) is relatively high, it suggests a strong deviation from what would be expected under no change. This indicates that patient admissions significantly changed between the two periods, likely due to the Ibis intervention.

```{r}
# Logistic Regression: Predict probability of admission.
logit_model <- glm(ibis_inpatient ~ pre_ibis_inpatient, data = events_full_new, family = "binomial")
summary(logit_model)
```

```{r}
inpatient_table <- pre_ibis_inpatient_table %>% 
  full_join()
```

```{r}
ibis_yes_no <- inpatient_table %>% 
  pivot_longer(c(3,4), names_to = "pre_post", values_to = "admit_count")
%>% 
  mutate(admit = if_else(admit_count == 0,0,1))
```

```{r}
glm(admit ~ pre_post, ibis_yes_no, family = "binomial") %>% summary()
 

```

Test
